---
title: "Bias-Variance Practical Assignment"
author: "José Pinto, Nirbhaya Shaji"
date: "28/02/2020"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intruduction

The bias-variance trade off is an important characteristic of machine learning processes that has to be
considered when making decisions about a learning task. Poor decisions can lead to poor a generalization
ability of the models and a failure of the learning task. It is important to know how methods and models
behave with different parameter values. Expressive models have a high bias and low variance. Less expressive
models have low bias and high variance.

## Goals

The goal of this assignment is to study the bias-variance trade off and expressivness of diferent models.<br>
In order to complete that goal we must train and test a variety of regression models, plot and describe the
training and test error and their evolution as we change parameters. We need to figure out wich parameters
increase the model expressivness and how they change it.

## Models

For this assignment a variety of models will be used.<br>
The models are <b>kNN, linear regression, LDA, QDA, Logistic regression, decision trees, polinomial regression</b> and <b>random forests</b>.<br> 
The models have different levels of complexity, for this analisys the models ca be seen as having low, medium an high complexity.<br>
Models with <b>low</b> complexity are <b>knn, linear regression</b> and <b>decision trees</b>.<br>
Models with <b>meduim</b> complexity are <b>LDA, QDA</b> and <b>Logistic regression</b>.<br>
Models with <b>high</b> complexity are <b>polinomial regression</b> and <b>random forests.</b><br>

## Dataset

The dataset we are going to work with is the "Pima Indians Diabetes" dataset available in the "mlbench"
library. The dataset contains information about the presence of diabetes in a population.<br>

The dataset is comprised of a total of 768 observations (rows) and 9 variables (columns), 1 target and 8
predictors.<br>
The variables are as follows:<br>
Target variable:<br>
<b>diabetes</b> - Presence of abcense of diabetes (test for diabetes) (binary)<br>

predictor variables:<br>
<b>pregnant</b> - Number of times pregnant(integer value)<br>
<b>glucose</b>	- Plasma glucose concentration (glucose tolerance test) (integer value)<br>
<b>pressure</b> - Diastolic blood pressure (mm Hg) (integer value)<br>
<b>triceps</b> -	Triceps skin fold thickness (mm) (integer value))<br>
<b>insulin</b> -	2-Hour serum insulin (mu U/ml) (integer value)<br>
<b>mass</b> - Body mass index (weight in kg/(height in m)\^2) (decimal value)<br>
<b>pedigree</b> - Diabetes pedigree function (decimal value)<br>
<b>age</b> - Age (years) (integer value)<br>

Due to the nature of the target this is a classification problem where we predict
"neg" or "pos", negative or positive diabetes.<br>

All the required libaries are imported here to more easily identify dependencies.<br>

```{r libraries, message=FALSE, warning=FALSE}
library(mlbench)
library(class)
library(GGally)
library(magrittr)
library(MASS)
library(hmeasure)
library(randomForest)
library(reshape2)
library(glmnet)
library(ggplot2)
library(reshape2)
```

Import the data and check variable types.<br>

```{r dataset}
data(PimaIndiansDiabetes)
str(PimaIndiansDiabetes)
```

First entries of the data.<br>

```{r head}
head(PimaIndiansDiabetes)
```

Summary of data distribution.<br>

```{r summary}
summary(PimaIndiansDiabetes)
```

Scatterplots, histograms, boxplots and correlation of data.<br>

```{r pairs}
PimaIndiansDiabetes %>% ggpairs(.,mapping=ggplot2::aes(colour=diabetes));
```

From the plots and summarys we spot no strong multicollinarity.<br>

```{r seed}
#set the seed so we can replicate results
set.seed(123)
```
## Data Split

Now that we have the data we split it into train and test data.<br>

```{r dataSplit}
#set the target column
targetColumn = 9
#set percentage of data for training to 80%
trainPercent = 0.8
#get the amount of rows for training
trainSize = floor(trainPercent*nrow(PimaIndiansDiabetes))
#get the indices of the rows for training
trainRows = sample(nrow(PimaIndiansDiabetes),trainSize,replace = FALSE)
#get the rows with the training indices
trainDiabetes = PimaIndiansDiabetes[trainRows,-targetColumn]
trainDiabetesY = PimaIndiansDiabetes[trainRows,targetColumn]
#get the rows that are not training indices (test data)
testDiabetes = PimaIndiansDiabetes[-trainRows,-targetColumn]
testDiabetesY = PimaIndiansDiabetes[-trainRows,targetColumn]
```

## Knn

The knn method creates a model with all points in the training data.<br>
For prediction it finds what are the k nearest neighbours and returns the most common
class.<br>
The value k of neighbours is the only hyperparameter to tune. It is expected that with
the increase in k the bias will increase and the variance will decrease, meaning as k
increases the expressivness will decrease.<br>
The model is capable of both the highest and lowest expressivness of all models tested,
with k equal to the number of examples in the training data it predicting no more than
the mode (the most common value in the whole data).<br>

```{r knn}
error = data.frame()

#get models, predictions and error for different values of k
for(k in seq(1,100,2)){
  trainDiabetesPred = knn(trainDiabetes,trainDiabetes,trainDiabetesY,k=k)
  testDiabetesPred = knn(trainDiabetes,testDiabetes,trainDiabetesY,k=k)
  trainError = 1 - sum(trainDiabetesPred==trainDiabetesY)/length(trainDiabetesY)
  testError = 1 - sum(testDiabetesPred==testDiabetesY)/length(testDiabetesY)
  
  error = rbind(error,data.frame(k,trainError,testError))
}

#get the maximum value of error (to set the max y value in the plot)
maxY = max(error$trainError,error$testError)

#create the empty plot with some parameters
plot(error[,1],main="Evolution of the train and test errors",xlab="Number of NN (k)",ylab="Error",ylim = c(0,maxY),type="n",xaxt="n")
#set the axis for the number of neighbours (k)
axis(1, 1:length(error[,1]), error[,1])
#plot the error for the training data
lines(error[,2],type="b",col='red',pch=20)
#plot the error for the testing data
lines(error[,3],type="b",col='blue',pch=20)

legend("bottomright",lty=1,col=c("red","blue"),legend = c("Train ", "Test "))
```

We can see in the graph above the confirmation of what was stated before, with
the training error increasing and testing error decreasing as we increase k.<br>

From now on we will use the value 13 for k as it has a good train and test error.<br>

## Decision boundaries

```{r knnb}
boundaryKNN <- function(X, Y, k, resolution = 100, newData = NULL, newDataY = NULL) {
  if(is.null(newData)){
    newData = X
    newDataY = Y
  }
  
  #get ranes for the data (min and max)
  ranges <- sapply(newData[,c(1:2)], range)
  #get n points bettwen the ranges
  xAxis <- seq(ranges[1,1], ranges[2,1], length.out = resolution)
  yAxis <- seq(ranges[1,2], ranges[2,2], length.out = resolution)
  #create a grid of points
  grid <- data.frame(expand.grid(xAxis,yAxis))
  
  #get the predicted values for the grid
  pred <- knn(X, grid, Y, k=k)
  #get the values as 0 and 1
  numericPred = as.numeric(pred)-1
  
  plotz<-ggplot(data=grid, aes(grid[,1], grid[,2])) +
    #place background color for predicions
    geom_point(aes(colour=pred), size=4,shape=16,alpha=.02) +
    #place prediction borders
    geom_contour(aes(z=numericPred), colour='black', size = 1,alpha=1) +
    #place data points
    geom_point(data=newData, aes(newData[,1], newData[,2], colour=newDataY),size=3,shape=16,alpha=.8)+
    #select display colors
    scale_color_manual(values=c("#A4A4A4", "#E61616")) +
    theme_bw() +
    #set x and y labels
    xlab(names(X)[1]) +
    ylab(names(X)[2])

  return(plotz)
}

```


Now we will present the decision boundaries for the model.<br>
First the boundaries for subsets of two atributes.<br>
From all the possible subsets of two variables two where selected.<br>

The first one shows good separation of the data and not a lot of overfitting.<br>

```{r knnBoundary1}
k=13
print(boundaryKNN(trainDiabetes[,c("glucose","mass")],trainDiabetesY,k))
```

The second one creates an interesting pattern, with vertical lines.<br>
This is probably due to the high amount of insulin with value 0.<br>

```{r knnBoundary2}
print(boundaryKNN(trainDiabetes[,c("insulin","pedigree")],trainDiabetesY,k))
```


Now we are going to draw the decision boundaries obtained by the first two components of
PCA (principal component analisis).<br>

First we obtain the PCA transformation of the data and change it for plotting.<br>

```{r knnBoundaryPCASetup}
#get number of columns
ncols = length(trainDiabetes)

#get the principal components
diabetesPCA = prcomp(trainDiabetes, center = TRUE,scale. = TRUE)
#transform the training data into its principal components
diabetesPC = predict(diabetesPCA,trainDiabetes)[,1:2]
```

Then we create a grid in the full dimentional space, make predictions in it and transform
it into the PCA space.<br>

```{r knnBoundaryPCAPredictGrid}
#create axis for the grid in the full dimentional space
gridAxis = c()
for(column in seq(1,ncols,1)){
  gridAxis = c(
    gridAxis,
    #get n equally spaced values bettwen the minimum of the column and the maximum 
    list(seq(
      min(trainDiabetes[,column]),
      max(trainDiabetes[,column]),
      #number n of values (low due to high dimentionality)
      length.out = 3
    ))
  )
}
#create a grid with all combinations of all axis
grid = data.frame(as.matrix(expand.grid(gridAxis)))
#set column names to names of training data(required to convert to pca)
names(grid) = names(trainDiabetes)
#convert grid values to pca space
gridPCA = predict(diabetesPCA,grid)[,1:2]

#get the predicted values for each element of the grid
pred = factor(knn(trainDiabetes,grid,trainDiabetesY,k=k))
levels(pred) = c("neg","pos")
#plot the contours and add the points
boundaryKNN(
  data.frame(gridPCA),
  pred,
  k=1,
  newData=data.frame(diabetesPC),
  newDataY=trainDiabetesY
)

```

## Linear regression 

Change the categorical target to numerical (neg = 0 and pos = 1).<br>

```{r LinearRegressionPreprocess}
trainDiabetesY_num = as.numeric(trainDiabetesY) - 1
testDiabetesY_num = as.numeric(testDiabetesY) - 1
```

linear model<br>

```{r LinearRegressionModel}
lm = lm(trainDiabetesY_num~.,data = trainDiabetes)
summary(lm)
```

R-squared:  0.2883 > not so good. remove NA may be?<br>

```{r LinearRegressionPredictions}
trainDiabetesPred = predict(lm,trainDiabetes)
tableTrain = table(trainDiabetesY_num,trainDiabetesPred>0.5)
tableTrain

testDiabetesPred = predict(lm,testDiabetes)
tableTest = table(testDiabetesY_num,testDiabetesPred>0.5)
tableTest
```

```{r LinearRegressionError}
trainError = 1 - sum(diag(tableTrain))/sum(tableTrain)
trainError

testError = 1 - sum(diag(tableTest))/sum(tableTest)
testError

error = data.frame()
error = rbind(error,data.frame(trainError,testError))
error
```

```{r LinearRegressionAccuracy}
lscm <- function(regmodel, observation){
  val <- predict.lm(regmodel,observation)
  sapply(val,function(v) if(v<0.5) 0 else 1 )
}

# classes obtained with linear regression as 0 and 1 (test)
testPred_binary = lscm(lm, data.frame(y = NA, testDiabetes))
accuracytest<- sum(testPred_binary == testDiabetes)/nrow(testDiabetes)*100 # 
accuracytest
```

## Polynomial regression 

```{r Polynomial Regression}
y = trainDiabetesY_num
trainx = data.matrix(trainDiabetes, rownames.force = NA)
testx = data.matrix(testDiabetes, rownames.force = NA)

#Squared

trainxSquare = trainx^2
testxSquare = testx^2

lm.Square = lm(y ~ .,data=data.frame(trainx,trainxSquare))
summary(lm.Square)
# R2 0.344 still not so good! but better than simple linear

trainDiabetesPredSquare =  predict(lm.Square,data.frame(trainx,trainxSquare))
tableTrainSquare = table(trainDiabetesY_num,trainDiabetesPredSquare>0.5)
tableTrainSquare
# trainDiabetesY_num FALSE TRUE
# 0   366   35
# 1    94  119
trainErrorSquare = 1 - sum(diag(tableTrainSquare))/sum(tableTrainSquare)
trainErrorSquare
# 0.2100977

toClass <- function(trainDiabetesPredSquare){
  val <- trainDiabetesPredSquare
  sapply(val,function(v) if(v<0.5) 0 else 1 )
}

trainDiabetesPredSquare_toclass = as.numeric(toClass(trainDiabetesPredSquare))

testDiabetesPredSquare = predict(lm.Square,data.frame(testx,testxSquare))
tableTestSquare = table(testDiabetesY_num,testDiabetesPredSquare>0.5)
tableTestSquare
# testDiabetesY_num FALSE TRUE
# 0    90    9
# 1    25   30
testErrorSquare = 1 - sum(diag(tableTestSquare))/sum(tableTestSquare)
testErrorSquare
# [1] 0.2207792



testPredSquare_binary = lscm(lm.Square, data.frame(y = NA, data.frame(testx,testxSquare))) # classes obtained with linear regression as 0 and 1 (test)
accuracytestSquare = sum(testPredSquare_binary == testDiabetesY_num)/nrow(testDiabetes)*100 
accuracytestSquare
# [1] 77.92208 
x = misclassCounts(testPredSquare_binary,testDiabetesY_num)
x$metrics$Sens
# [1] 0.5454545


# ER      Sens      Spec Precision    Recall       TPR        FPR         F    Youden
# 1 0.2207792 0.5454545 0.9090909 0.7692308 0.5454545 0.5454545 0.09090909 0.6382979 0.4545455

##Cubic

trainxCubic = trainx^3
testxCubic = testx^3

lm.Cubic = lm(y ~ .,data=data.frame(trainx,trainxSquare,trainxCubic))
summary(lm.Cubic)
# R2 0.3567 still not so good but better than Square

trainPredCubic_binary_toClass = lscm(lm.Cubic, data.frame(y = NA, data.frame(trainx,trainxSquare,trainxCubic)))
trainDiabetesPredCubic =  predict(lm.Cubic,data.frame(trainx,trainxSquare,trainxCubic))
tableTrainCubic = table(trainDiabetesY_num,trainDiabetesPredCubic>0.5)
tableTrainCubic
# trainDiabetesY_num FALSE TRUE
# 0   351   50
# 1    89  124
trainErrorCubic = 1 - sum(diag(tableTrainCubic))/sum(tableTrainCubic)
trainErrorCubic
# [1] 0.2263844


testDiabetesPredCubic = predict(lm.Cubic,data.frame(testx,testxSquare,testxCubic))
tableTestCubic = table(testDiabetesY_num,testDiabetesPredCubic>0.5)
tableTestCubic
# testDiabetesY_num FALSE TRUE
# 0    88   11
# 1    25   30
testErrorCubic = 1 - sum(diag(tableTestCubic))/sum(tableTestCubic)
testErrorCubic
# [1] 0.2337662 


testPredCubic_binary = lscm(lm.Cubic, data.frame(y = NA, data.frame(testx,testxSquare,testxCubic))) # classes obtained with linear regression as 0 and 1 (test)
accuracytestCubic = sum(testPredCubic_binary == testDiabetesY_num)/nrow(testDiabetes)*100 
accuracytestCubic
# [1] 76.62338
x = misclassCounts(testPredCubic_binary,testDiabetesY_num)
x$metrics$Sens
#0.5454545


# ER      Sens      Spec Precision    Recall       TPR       FPR     F    Youden
# 1 0.2337662 0.5454545 0.8888889 0.7317073 0.5454545 0.5454545 0.1111111 0.625 0.4343434

##Quartic



trainxQuartic = trainx^4
testxQuartic = testx^4

lm.Quartic = lm(y ~ .,data=data.frame(trainx,trainxSquare,trainxCubic,trainxQuartic))
summary(lm.Quartic)
# R2 0.3649 still not so good but better than Square and Cubic

testPredQuartic_binary_toClass = lscm(lm.Quartic, data.frame(y = NA, data.frame(trainx,trainxSquare,trainxCubic,trainxQuartic)))
trainDiabetesPredQuartic =  predict(lm.Quartic,data.frame(trainx,trainxSquare,trainxCubic,trainxQuartic))
tableTrainQuartic = table(trainDiabetesY_num,trainDiabetesPredQuartic>0.5)
tableTrainQuartic
# trainDiabetesY_num FALSE TRUE
# 0   356   45
# 1    86  127
trainErrorQuartic = 1 - sum(diag(tableTrainQuartic))/sum(tableTrainQuartic)
trainErrorQuartic
# 0.213355


testDiabetesPredQuartic = predict(lm.Quartic,data.frame(testx,testxSquare,testxCubic,testxQuartic))
tableTestQuartic = table(testDiabetesY_num,testDiabetesPredQuartic>0.5)
tableTestQuartic
# testDiabetesY_num FALSE TRUE
# 0    86   13
# 1    23   32
testErrorQuartic = 1 - sum(diag(tableTestQuartic))/sum(tableTestQuartic)
testErrorQuartic
# [1] 0.2337662 same as cubic!


testPredQuartic_binary = lscm(lm.Quartic, data.frame(y = NA, data.frame(testx,testxSquare,testxCubic,testxQuartic))) # classes obtained with linear regression as 0 and 1 (test)
accuracytestQuartic = sum(testPredQuartic_binary == testDiabetes)/nrow(testDiabetes)*100 
accuracytestQuartic
# [1] 68.83117 #least accurate
x = misclassCounts(testPredQuartic_binary,testDiabetesY_num)
x$metrics$Sens
# 0.5818182



##quintic


trainxQuintic = trainx^5
testxQuintic = testx^5

lm.Quintic = lm(y ~ .,data=data.frame(trainx,trainxSquare,trainxCubic,trainxQuartic,trainxQuintic))
summary(lm.Quintic)
# R2 0.3692 still not so good but better than Square Cubic and Quartic 

testPredQuintic_binary_toClass = lscm(lm.Quintic, data.frame(y = NA, data.frame(trainx,trainxSquare,trainxCubic,trainxQuartic,trainxQuintic)))
  trainDiabetesPredQuintic =  predict(lm.Quintic,data.frame(trainx,trainxSquare,trainxCubic,trainxQuartic,trainxQuintic))
tableTrainQuintic = table(trainDiabetesY_num,trainDiabetesPredQuintic>0.5)
tableTrainQuintic
# trainDiabetesY_num FALSE TRUE
# 0   353   48
# 1    89  124
trainErrorQuintic = 1 - sum(diag(tableTrainQuintic))/sum(tableTrainQuintic)
trainErrorQuintic
# 0.223127



testDiabetesPredQuintic = predict(lm.Quintic,data.frame(testx,testxSquare,testxCubic,testxQuartic,testxQuintic))
tableTestQuintic = table(testDiabetesY_num,testDiabetesPredQuintic>0.5)
tableTestQuintic
# testDiabetesY_num FALSE TRUE
# 0    86   13
# 1    24   31
testErrorQuintic = 1 - sum(diag(tableTestQuintic))/sum(tableTestQuintic)
testErrorQuintic
# [1] [1] 0.2402597



testPredQuintic_binary = lscm(lm.Quintic, data.frame(y = NA, data.frame(testx,testxSquare,testxCubic,testxQuartic,testxQuintic))) # classes obtained with linear regression as 0 and 1 (test)
accuracytestQuintic = sum(testPredQuintic_binary == testDiabetes)/nrow(testDiabetes)*100 
accuracytestQuintic
# [1] 70.77922 
x = misclassCounts(testPredQuintic_binary,testDiabetesY_num)
x$metrics$Sens
# 0.5636364

# ER      Sens      Spec Precision    Recall       TPR       FPR         F    Youden
# 1 0.2402597 0.5636364 0.8686869 0.7045455 0.5636364 0.5636364 0.1313131 0.6262626 0.4323232

```

## Logistic Regression

```{r Logistic Regression}

glm_model = glm(trainDiabetesY_num~.,data = trainDiabetes, family = binomial)
summary(glm_model)

glm_prob = predict.glm(glm_model,trainDiabetes,type='response')
glm_predict = rep('neg',nrow(trainDiabetes))
glm_predict[glm_prob>.5] = 'pos'
glmTrain_table = table(pred=glm_predict,true=trainDiabetesY)
glmTrain_table
# pred  neg pos
# neg 354  90
# pos  47 123
glm_trainError = 1 - sum(diag(glmTrain_table))/sum(glmTrain_table)
glm_trainError
# [1] 0.223127



glm_prob = predict.glm(glm_model,testDiabetes,type='response')
contrasts(PimaIndiansDiabetes$diabetes)
glm_predict = rep('neg',nrow(testDiabetes))
glm_predict[glm_prob>.5] = 'pos'
glmTest_table = table(pred=glm_predict,true=testDiabetesY)
glmTest_table
# pred  neg pos
# neg  87  24
# pos  12  31
glm_testError = 1 - sum(diag(glmTest_table))/sum(glmTest_table)
glm_testError
# 0.2337662



glmAccuracy = mean(glm_predict==testDiabetesY)*100
glmAccuracy
# 76.62338
x = misclassCounts(as.numeric(as.factor(glm_predict))-1,testDiabetesY_num)
x$metrics$Sens
# [1] 0.5636364



# ER      Sens      Spec Precision    Recall       TPR       FPR         F    Youden
# 1 0.2337662 0.5636364 0.8787879 0.7209302 0.5636364 0.5636364 0.1212121 0.6326531 0.4424242
```

```{r Lasso}

#input the predictors as a matrix and the class labels as a vector (numeric)

x = model.matrix(trainDiabetesY_num~.,trainDiabetes)
y = trainDiabetesY_num
set.seed(2322111)
cv.out = cv.glmnet(x,y,alpha=1,family='binomial',type.measure = 'mse' )
plot(cv.out)

#min value of lambda
lambda_min = cv.out$lambda.min
#best value of lambda
lambda_1se = cv.out$lambda.1se
#regression coefficients
coef(cv.out,s=lambda_1se)
# 10 x 1 sparse Matrix of class "dgCMatrix"
# 1
# (Intercept) -5.277686156
# (Intercept)  .          
# pregnant     0.065718855
# glucose      0.024448506
# pressure     .          
# triceps      .          
# insulin      .          
# mass         0.034426079
# pedigree     0.118308012
# age          0.005109729


x_train = model.matrix(trainDiabetesY_num~.,trainDiabetes)

lasso_probtrain <- predict(cv.out,newx = x_train,s=lambda_1se,type='response')
lasso_predicttrain <- rep('neg',nrow(trainDiabetes))
lasso_predicttrain[lasso_probtrain>.5] <- 'pos'
lassoTrain_table = table(pred=lasso_predicttrain,true=trainDiabetesY_num)
lasso_trainError = 1 - sum(diag(lassoTrain_table))/sum(lassoTrain_table)
lasso_trainError
# 0.2263844


#get test data
x_test = model.matrix(testDiabetesY_num~.,testDiabetes)
#predict class, type=”class”
lasso_prob <- predict(cv.out,newx = x_test,s=lambda_1se,type='response')
#translate probabilities to predictions
lasso_predict <- rep('neg',nrow(testDiabetes))
lasso_predict[lasso_prob>.5] <- 'pos'
#confusion matrix
lassoTest_table = table(pred=lasso_predict,true=testDiabetesY_num)
# red   0  1
# neg 89 28
# pos 10 27
#accuracy
 

# pred   0  1
# neg 90 26
# pos  9 29

lasso_testError = 1 - sum(diag(lassoTest_table))/sum(lassoTest_table)
lasso_testError
# 0.2467532


mean(lasso_predict==testDiabetesY)
#[1] 0.7532468
x = misclassCounts(as.numeric(as.factor(lasso_predict))-1,testDiabetesY_num)
x$metrics$Sens
# 0.4909091

```

```{r Lasso + Ridge - Elastic net}
########## elstic net (Ridge + Lasso)

x = model.matrix(trainDiabetesY_num~.,trainDiabetes)
y = trainDiabetesY_num
set.seed(2322111)
cv.out = cv.glmnet(x,y,alpha=1,family='binomial',type.measure = 'mse', relax = TRUE)
plot(cv.out)

#min value of lambda
lambda_min = cv.out$lambda.min
#best value of lambda
lambda_1se = cv.out$lambda.1se
#regression coefficients
coef(cv.out,s=lambda_1se)
# 10 x 1 sparse Matrix of class "dgCMatrix"
# 1
# (Intercept) -5.277686156
# (Intercept)  .          
# pregnant     0.065718855
# glucose      0.024448506
# pressure     .          
# triceps      .          
# insulin      .          
# mass         0.034426079
# pedigree     0.118308012
# age          0.005109729

x_train = model.matrix(trainDiabetesY_num~.,trainDiabetes)
lasso_probtrain <- predict(cv.out,newx = x_train,s=lambda_1se,type='response')
lasso_predicttrain <- rep('neg',nrow(trainDiabetes))
lasso_predicttrain[lasso_probtrain>.5] <- 'pos'
lassoTrain_table = table(pred=lasso_predicttrain,true=trainDiabetesY_num)
lasso_trainError = 1 - sum(diag(lassoTrain_table))/sum(lassoTrain_table)
lasso_trainError
# 0.223127



#get test data
x_test = model.matrix(testDiabetesY_num~.,testDiabetes)
#predict class, type=”class”
lasso_prob <- predict(cv.out,newx = x_test,s=lambda_1se,type='response')
#translate probabilities to predictions
lasso_predictEla <- rep('neg',nrow(testDiabetes))
lasso_predictEla[lasso_prob>.5] <- 'pos'
#confusion matrix
lassoTest_table =  table(pred=lasso_predictEla,true=testDiabetesY_num)

# pred   0  1
# neg 90 26
# pos  9 29

lasso_testError = 1 - sum(diag(lassoTest_table))/sum(lassoTest_table)
lasso_testError
# 0.2402597 


#accuracy
mean(lasso_predictEla==testDiabetesY)
#[1] 0.7597403
misclassCounts(as.numeric(as.factor(lasso_predictEla))-1,testDiabetesY_num)
#Sensitivity 0.5272727

```

```{r Sensitivity and Error Plots}

predictionsName = c('testPred_binary','testPredSquare_binary','testPredCubic_binary','testPredQuartic_binary',
                 'testPredQuintic_binary','glm_predict','lasso_predict','lasso_predictEla')


predFrame = as.data.frame(matrix(nrow=length(testDiabetesY_num)),ncol=0)
predFrame$testPred_binary = testPred_binary
predFrame$testPredSquare_binary = testPredSquare_binary
predFrame$testPredCubic_binary = testPredCubic_binary
predFrame$testPredQuartic_binary = testPredQuartic_binary
predFrame$testPredQuintic_binary = testPredQuintic_binary
#predFrame$glm_predict = glm_predict
predFrame$glm_predict = as.numeric(as.factor(glm_predict))-1
predFrame$lasso_predict = as.numeric(as.factor(lasso_predict))-1
predFrame$lasso_predictEla = as.numeric(as.factor(lasso_predictEla))-1
predFrame = predFrame[,-1]

metrics = data.frame()

for(i in 1:ncol(predFrame)){
  
    metrics = rbind(metrics,as.data.frame(misclassCounts(predFrame[,i],testDiabetesY_num)$metrics))
  
}


metrics$model = c('Liner Regretion','2nd Order Polynomial Regression','3rd Order Polynomial Regression','4th Order Polynomial Regression',
                  '5th Order Polynomial Regression','Logistic Regression','Lasso','Elastic Net (Lasso + Ridge)')


ggplot(data=metrics, aes(x = model , y = Sens, col=model))+ geom_point(aes(size = Sens)) + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank()) + labs(y = "Sensitivity")

errordt = data.frame()
errordt = data.frame(train = c(0.2263844,0.2100977,0.2263844,0.213355,0.223127,0.223127,0.2263844,0.223127), test = c(0.2272727,0.2207792,0.2337662,0.2337662,0.2402597,0.2337662,0.2467532,0.2402597), model = c('Liner Regretion','2nd Order Polynomial Regression','3rd Order Polynomial Regression','4th Order Polynomial Regression',
                                                                                                                      '5th Order Polynomial Regression','Logistic Regression','Lasso','Elastic Net (Lasso + Ridge)') )
str(errordt)

errordt = melt(errordt)
ggplot(errordt, aes(variable, value, group=factor(model))) + geom_point(aes(color=factor(model)))

ggplot(errordt, aes(variable, value, group=factor(model))) + geom_line(aes(color=factor(model),linetype=factor(model))) 

```

```{r Decision Boundry Nirbhaya}
#So many. need to make some choices.
```

```{r LDA}
PimaTrain<-trainDiabetes
PimaTrain$diabetes<-trainDiabetesY
PimaTest<-testDiabetes
PimaTest$diabetes<-testDiabetesY

lda.pima<-lda(diabetes~., data = PimaTrain)
pred.ldatest.pima<-predict(lda.pima,PimaTest)$class
pred.ldatrain.pima<-predict(lda.pima,PimaTrain)$class

#QDA
qda.pima<-qda(diabetes~., data = PimaTrain)
pred.qdatest.pima<-predict(qda.pima,PimaTest)$class
pred.qdatrain.pima<-predict(qda.pima,PimaTrain)$class


#comparing stuff
ldaqda=rbind(misclassCounts(as.numeric(pred.ldatest.pima)-1,as.numeric(testDiabetesY)-1)$metrics,
             misclassCounts(as.numeric(pred.ldatrain.pima)-1,as.numeric(trainDiabetesY)-1)$metrics,
             misclassCounts(as.numeric(pred.qdatest.pima)-1,as.numeric(testDiabetesY)-1)$metrics,
             misclassCounts(as.numeric(pred.qdatrain.pima)-1,as.numeric(trainDiabetesY)-1)$metrics)
ldaqda$model<-c("LDATest","LDATrain","QDATest","QDATrain")


ggplot()+geom_point(data=ldaqda,aes(x=model,y=Sens,colour=model))


```

```{r RandomForestNTree}
#TREE NUMBER

oob.err<-double(500)
ptest.err<-double(500)
ntest.err<-double(500)
test.err<-data.frame()
train.err<-data.frame()

for(ntree in seq(1,500)) {
  rf.model<-randomForest(diabetes ~ . , data = PimaTrain ,mtry=4,ntree=ntree) 
  oob.err[ntree] = rf.model$err.rate[ntree,1] 
  ptest.err[ntree]<-rf.model$err.rate[ntree,2]
  ntest.err[ntree]<-rf.model$err.rate[ntree,3]
  
  pred<-predict(rf.model,PimaTest) 
  predTrain<-predict(rf.model,PimaTrain)
  test.err<- rbind(test.err,as.data.frame(misclassCounts(as.numeric(pred)-1,as.numeric(testDiabetesY)-1)$metrics))
  train.err<- rbind(train.err,as.data.frame(misclassCounts(as.numeric(predTrain)-1,as.numeric(trainDiabetesY)-1)$metrics))
  
  
  #cat(ntree," ") 
  
}


#OOB, negative, positive in function of mtry
plot(ntest.err,type="l",ylim=c(0.15, 0.5),xlim=c(0,500))
lines(ptest.err,col="red")
lines(oob.err,col="green")


#performance evaluatores in function of mtry
test.err$id = row.names(test.err)
test.err.long<-melt(test.err,id.vars="id")

train.err$id=row.names(train.err)
train.err.long<-melt(train.err,id.vars="id")
ggplot()+geom_line(data=test.err.long,aes(x=as.numeric(id),y=value,colour=variable,group=variable),size=1)+xlab("ntree")+ylab("Criteria")

ggplot()+geom_line(data=test.err,aes(x=as.numeric(id),y=Sens))+geom_line(data=train.err,aes(x=as.numeric(id),y=Sens),col="red")+xlab("Trees")
```


```{r RandomForestMaxPred}
#PREDICTORS PER NODE

test.err<-data.frame()
train.err<-data.frame()



for(mtry in 1:8) {
  rf.model<-randomForest(diabetes ~ . , data = PimaTrain ,mtry=mtry,ntree=200) 
  
  pred<-predict(rf.model,PimaTest) 
  predTrain<-predict(rf.model,PimaTrain)
  test.err<- rbind(test.err,as.data.frame(misclassCounts(as.numeric(pred)-1,as.numeric(testDiabetesY)-1)$metrics))
  train.err<- rbind(train.err,as.data.frame(misclassCounts(as.numeric(predTrain)-1,as.numeric(trainDiabetesY)-1)$metrics))
  
  cat(mtry," ")

}

test.err$id = row.names(test.err)
train.err$id=row.names(train.err)
test.err.long<-melt(test.err,id.vars="id")
ggplot()+geom_line(data=test.err.long,aes(x=as.numeric(id),y=value,colour=variable,group=variable),size=1)+xlab("Predictors per Split")+ylab("Criteria")

ggplot()+geom_line(data=test.err,aes(x=as.numeric(id),y=Sens))+geom_line(data=train.err,aes(x=as.numeric(id),y=Sens),col="red")+xlab("Predictors p/ Node")
```


```{r RandomForestMinNode}
#NODESIZE

test.err<-data.frame()
train.err<-data.frame()

for(nodesize in seq(1,40)) {
  rf.model<-randomForest(diabetes ~ . , data = PimaTrain ,mtry=4,ntree=200,nodesize=nodesize) 
  pred<-predict(rf.model,PimaTest) 
  predTrain<-predict(rf.model,PimaTrain)
  test.err<- rbind(test.err,as.data.frame(misclassCounts(as.numeric(pred)-1,as.numeric(testDiabetesY)-1)$metrics))
  train.err<- rbind(train.err,as.data.frame(misclassCounts(as.numeric(predTrain)-1,as.numeric(trainDiabetesY)-1)$metrics))
  
  
  cat(nodesize," ") 
  
}

test.err$id = row.names(test.err)
train.err$id=row.names(train.err)
test.err.long<-melt(test.err,id.vars="id")

ggplot()+geom_line(data=test.err.long,aes(x=as.numeric(id),y=value,colour=variable,group=variable),size=1)+xlab("Node Size")+ylab("Criteria")

ggplot()+geom_line(data=test.err,aes(x=as.numeric(id),y=Sens))+geom_line(data=train.err,aes(x=as.numeric(id),y=Sens),col="red")+xlab("Node Size")

```

```{r Boundary plot function, works for a few models}
boundaryplot <- function(model, data1, class = NULL, predict_type = "class",
                             rezz = 150, showgrid = TRUE, ...) {
  
  if(!is.null(class)) cl <- data1[,class] else cl <- 1
  data1 <- data1[,1:2]
  
  
  
  r <- sapply(data1[,c(1:2)], range, na.rm = TRUE)
  xs <- seq(r[1,1], r[2,1], length.out = rezz)
  ys <- seq(r[1,2], r[2,2], length.out = rezz)
  g <- cbind(rep(xs, each=rezz), rep(ys, time = rezz))
  colnames(g) <- colnames(r)
  g <- as.data.frame(g)
  
  #if(is.list(model)) {
  p <- predict(model, g)
  if(is.list(p)) p <- p$class
  p <- as.factor(p)
  g$diabetes<-p
  z <- matrix(as.integer(p), nrow = rezz, byrow = TRUE)
  #}
  #else {g$diabetes<-model} 
  g$z <- c(0, 1)[sapply(g$diabetes, as.numeric)]
  
  data1$diabetes<-cl
  
  plotz<-ggplot(data=g, aes(g[,1], g[,2])) +
    geom_point(aes(colour = diabetes), size=4,shape=16,alpha=.02) +
    geom_contour(aes(z=z), colour='black', size = 1,alpha=1) +
    geom_point(data=data1, aes(data1[,1], data1[,2], colour=data1[,3]),size=3,shape=16,alpha=.8) +
    scale_color_manual(values=c("#A4A4A4", "#E61616")) +
    theme_bw() +
    xlab(names(data1)[1]) +
    ylab(names(data1)[2])
  
  return(plotz)
  
}
```


```{r pre-processing just for fun}
#mean imputation, normalize, PCA

PimaIndiansDiabetes[PimaIndiansDiabetes==0]<-NA


NA2mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
dftest<-replace(PimaIndiansDiabetes, TRUE, lapply(PimaIndiansDiabetes, NA2mean))

normalize1 <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}


for(i in 1:8){
  dftest[,i]<-normalize1(dftest[,i])
} 

pima.pca.norm<-prcomp(dftest[,c(1:8)],center=TRUE,scale.=TRUE)

summary(pima.pca.norm)

pima.pca.norm.pred = predict(pima.pca.norm,dftest[,c(1:8)])

df8<-cbind(as.data.frame(pima.pca.norm.pred[,c(1,2)]),PimaIndiansDiabetes$diabetes)
names(df8)[names(df8)=="PimaIndiansDiabetes$diabetes"]<-"diabetes"

lda.df8<-lda(diabetes~.,data=df8)
qda.df8<-qda(diabetes~.,data=df8)
rf.df8<-randomForest(diabetes ~ . , data = df8,ntree=200,nodesize=12)




```


```{r LDA plot}
boundaryplot(lda.df8,df8,class="diabetes",main="LDA")
```


```{r QDA plot}

boundaryplot(qda.df8,df8,class="diabetes",main="QDA")
```


```{r RF plot}
boundaryplot(rf.df8,df8,class="diabetes",main="Random Forest")
```

```R
fit <- rpart(trainDiabetesY~., data = trainDiabetes, method = 'class')
```
# Decision Tree

```R
rpart.plot(fit, extra = )
```
```R
predict_unseen <-predict(fit, testDiabetes, type = 'class')
```


```R
table = table(testDiabetesY, predict_unseen)
```


```R
table
```
```R
accuracy_test <- sum(diag(table)) / sum(table)
```

```R
metrics_default = misclassCounts(as.numeric(as.factor(predict_unseen))-1,as.numeric(testDiabetesY)-1)
```


```R
metrics_default
```
#cp
```R
sensitivities = vector(mode = 'list')
error_cp = data.frame(matrix(ncol = 3, nrow = 0))
```


```R
cps = c(0, 0.01, 0.02, 0.03, 0.04, 0.05, 0.06, 0.07, 0.08, 0.09, 0.1)
```


```R
for (cp in seq(0,0.1,0.01)) {
    control = rpart.control(cp = cp)
    fit_cp <- rpart(trainDiabetesY~., data = trainDiabetes, method = 'class', control = control)
    predict_unseen_cp <-predict(fit_cp, testDiabetes, type = 'class')
    table = table(testDiabetesY, predict_unseen_cp)
    metrics_cp = misclassCounts(as.numeric(as.factor(predict_unseen_cp))-1,as.numeric(testDiabetesY)-1)
    sensitivities = append(sensitivities, metrics_cp$metrics$Sens)
    trainDiabetesPred_cp = predict(fit_cp, trainDiabetes, type = 'class')
    testDiabetesPred_cp = predict(fit_cp, testDiabetes, type = 'class')
    trainError_cp = 1 - sum(trainDiabetesPred_cp==trainDiabetesY)/length(trainDiabetesY)
    testError_cp = 1 - sum(testDiabetesPred_cp==testDiabetesY)/length(testDiabetesY)
    error_cp = rbind(error_cp,data.frame(cp,trainError_cp,testError_cp))
}
maxY = max(error_cp$trainError_cp, error_cp$testError_cp)
```


```R
sens <- do.call(rbind, Map(data.frame, cp_values = cps, sensitivities = sensitivities))
```


```R
ggplot(sens, aes(x = cp_values, y = sensitivities)) + geom_line() + geom_point() + labs(x = 'Value of cp', y = 'Sensitivity')
```
```R
#create the empty plot with some parameters
plot(error_cp[,1],main="Evolution of the train and test errors",xlab="Value of cp",ylab="Error",ylim = c(0,maxY),type="n",xaxt="n")
#set the axis for the number of values
axis(1, 1:length(error_cp[,1]), error_cp[,1])
#plot the error for the training data
lines(error_cp[,2],type="b",col='red',pch=20)
#plot the error for the testing data
lines(error_cp[,3],type="b",col='blue',pch=20)
legend("bottomright",lty=1,col=c("red","blue"),legend = c("Train ", "Test "))
```

```R
ncols = length(trainDiabetes)
cp = 0.02

for(column1 in seq(1,ncols-1,1)){
  #select a diferent second column that has not been a first columns
  for(column2 in seq(column1+1,ncols,1)){
    #get the values from the first column
    attribute1 = trainDiabetes[column1]
    #get the values from the second column
    attribute2 = trainDiabetes[column2]
    #turn the target variable into a number (for visualization purposes)
    target = as.numeric(trainDiabetesY)
    
    #create points in the x axis
    gridX = seq(
      min(attribute1),max(attribute1),
      (max(attribute1)-min(attribute1))/100
    )
    #create points in the y axis
    gridY = seq(
      min(attribute2),max(attribute2),
      (max(attribute2)-min(attribute2))/100
    )
    #create a grid with the points
    grid = as.matrix(expand.grid(gridX,gridY))
    #get grid height
    gridWidth = length(gridX)
    #get grid width
    gridHeight = length(gridY)
  
    #plot the datapoints
    plot(
      data.frame(attribute1,attribute2), #create a data frame with the two selected columns
      col=ifelse(target==1,"blue","red"), #set the colors in wich to draw the points (based on target value)
      pch=20,cex=1.5,cex.lab=1.4 #visualization parameters
    )
    
    #get the predicted values for each element of the grid
    control = rpart.control(cp = cp)
    fit_cp_2 <- rpart(trainDiabetes[,c(column1,column2)], data = trainDiabetes, method = 'class', control = control)
    pred = as.numeric(predict(fit_cp_2, trainDiabetes, type = 'class'))
    #add the grid points to the plot
    points(grid,pch='.',col=ifelse(pred==1,"blue","red"))
    #add the decision boundary to the plot
    contour(gridX,gridY,matrix(pred,gridWidth,gridHeight),levels=1.5, add=TRUE, d=FALSE, lty=1)
  }
}
```
#minsplit

```R
sens_minsplit  = vector(mode = 'list')
error_minsplit = data.frame(matrix(ncol = 3, nrow = 0))
```


```R
for (minsplit in seq(20,40,1)) {
    control = rpart.control(minsplit = minsplit)
    fit_minsplit <- rpart(trainDiabetesY~., data = trainDiabetes, method = 'class', control = control)
    predict_unseen_minsplit <-predict(fit_minsplit, testDiabetes, type = 'class')
    table = table(testDiabetesY, predict_unseen_minsplit)
    metrics_tuned = misclassCounts(as.numeric(as.factor(predict_unseen_minsplit))-1,as.numeric(testDiabetesY)-1)
    sens_minsplit = append(sens_minsplit, metrics_tuned$metrics$Sens)
    trainDiabetesPred_minsplit = predict(fit_minsplit, trainDiabetes, type = 'class')
    testDiabetesPred_minsplit = predict(fit_minsplit, testDiabetes, type = 'class')
    trainError_minsplit = 1 - sum(trainDiabetesPred_minsplit==trainDiabetesY)/length(trainDiabetesY)
    testError_minsplit = 1 - sum(testDiabetesPred_minsplit==testDiabetesY)/length(testDiabetesY)
    error_minsplit = rbind(error_minsplit,data.frame(minsplit,trainError_minsplit,testError_minsplit))
}
maxY_ms = max(error_minsplit$trainError_minsplit, error_minsplit$testError_minsplit)
```


```R
values_ms = c(20:40)
```


```R
ms <- do.call(rbind, Map(data.frame, minsplit_values = values_ms, sensitivities = sens_minsplit))
```


```R
ggplot(ms, aes(x = minsplit_values, y = sensitivities)) + geom_line() + geom_point() + labs(x = 'Value of minsplit', y = 'Sensitivity')
```
```R
#create the empty plot with some parameters
plot(error_minsplit[,1],main="Evolution of the train and test errors",xlab="Value of minsplit",ylab="Error",ylim = c(0,maxY_ms),type="n",xaxt="n")
#set the axis for the number of neighbours (k)
axis(1, 1:length(error_minsplit[,1]), error_minsplit[,1])
#plot the error for the training data
lines(error_minsplit[,2],type="b",col='red',pch=20)
#plot the error for the testing data
lines(error_minsplit[,3],type="b",col='blue',pch=20)
legend("bottomright",lty=1,col=c("red","blue"),legend = c("Train ", "Test "))
```
```R
ncols = length(trainDiabetes)
minsplit = 29

for(column1 in seq(1,ncols-1,1)){
  #select a diferent second column that has not been a first columns
  for(column2 in seq(column1+1,ncols,1)){
    #get the values from the first column
    attribute1 = trainDiabetes[column1]
    #get the values from the second column
    attribute2 = trainDiabetes[column2]
    #turn the target variable into a number (for visualization purposes)
    target = as.numeric(trainDiabetesY)
    
    #create points in the x axis
    gridX = seq(
      min(attribute1),max(attribute1),
      (max(attribute1)-min(attribute1))/100
    )
    #create points in the y axis
    gridY = seq(
      min(attribute2),max(attribute2),
      (max(attribute2)-min(attribute2))/100
    )
    #create a grid with the points
    grid = as.matrix(expand.grid(gridX,gridY))
    #get grid height
    gridWidth = length(gridX)
    #get grid width
    gridHeight = length(gridY)
  
    #plot the datapoints
    plot(
      data.frame(attribute1,attribute2), #create a data frame with the two selected columns
      col=ifelse(target==1,"blue","red"), #set the colors in wich to draw the points (based on target value)
      pch=20,cex=1.5,cex.lab=1.4 #visualization parameters
    )
    
    #get the predicted values for each element of the grid
    control = rpart.control(minsplit = minsplit)
    fit_minsplit_2 <- rpart(trainDiabetes[,c(column1,column2)], data = trainDiabetes, method = 'class', control = control)
    pred = as.numeric(predict(fit_minsplit_2, trainDiabetes, type = 'class'))
    #add the grid points to the plot
    points(grid,pch='.',col=ifelse(pred==1,"blue","red"))
    #add the decision boundary to the plot
    contour(gridX,gridY,matrix(pred,gridWidth,gridHeight),levels=1.5, add=TRUE, d=FALSE, lty=1)
  }
}
```
#minbucket
```R
sens_minbucket  = vector(mode = 'list')
error_minbucket = data.frame(matrix(ncol = 3, nrow = 0))
```


```R
for (minbucket in seq(20/3,40/3,1)) {
    control = rpart.control(minbucket = minbucket)
    fit_minbucket <- rpart(trainDiabetesY~., data = trainDiabetes, method = 'class', control = control)
    predict_unseen_minbucket <-predict(fit_minbucket, testDiabetes, type = 'class')
    table = table(testDiabetesY, predict_unseen_minbucket)
    metrics_tuned = misclassCounts(as.numeric(as.factor(predict_unseen_minbucket))-1,as.numeric(testDiabetesY)-1)
    sens_minbucket = append(sens_minbucket, metrics_tuned$metrics$Sens)
    trainDiabetesPred_minbucket = predict(fit_minbucket, trainDiabetes, type = 'class')
    testDiabetesPred_minbucket = predict(fit_minbucket, testDiabetes, type = 'class')
    trainError_minbucket = 1 - sum(trainDiabetesPred_minbucket==trainDiabetesY)/length(trainDiabetesY)
    testError_minbucket = 1 - sum(testDiabetesPred_minbucket==testDiabetesY)/length(testDiabetesY)
    error_minbucket = rbind(error_minbucket,data.frame(minbucket,trainError_minbucket,testError_minbucket))
}
maxY_mb = max(error_minbucket$trainError_minbucket, error_minbucket$testError_minbucket)
```


```R
values_mb = c(20/3:40/3)
```


```R
mb <- do.call(rbind, Map(data.frame, minbucket_values = values_mb, sensitivities = sens_minbucket))
```

    Warning message in mapply(FUN = f, ..., SIMPLIFY = FALSE):
    "longer argument not a multiple of length of shorter"


```R
ggplot(mb, aes(x = minbucket_values, y = sensitivities)) + geom_line() + geom_point() + labs(x = 'Value of minbucket', y = 'Sensitivity')
```
```R
#create the empty plot with some parameters
plot(error_minbucket[,1],main="Evolution of the train and test errors",xlab="Value of minbucket",ylab="Error",ylim = c(0,maxY_mb),type="n",xaxt="n")
#set the axis for the number of neighbours (k)
axis(1, 1:length(error_minbucket[,1]), error_minbucket[,1])
#plot the error for the training data
lines(error_minbucket[,2],type="b",col='red',pch=20)
#plot the error for the testing data
lines(error_minbucket[,3],type="b",col='blue',pch=20)
legend("bottomright",lty=1,col=c("red","blue"),legend = c("Train ", "Test "))
```
```R
ncols = length(trainDiabetes)
minbucket = 10.67

for(column1 in seq(1,ncols-1,1)){
  #select a diferent second column that has not been a first columns
  for(column2 in seq(column1+1,ncols,1)){
    #get the values from the first column
    attribute1 = trainDiabetes[column1]
    #get the values from the second column
    attribute2 = trainDiabetes[column2]
    #turn the target variable into a number (for visualization purposes)
    target = as.numeric(trainDiabetesY)
    
    #create points in the x axis
    gridX = seq(
      min(attribute1),max(attribute1),
      (max(attribute1)-min(attribute1))/100
    )
    #create points in the y axis
    gridY = seq(
      min(attribute2),max(attribute2),
      (max(attribute2)-min(attribute2))/100
    )
    #create a grid with the points
    grid = as.matrix(expand.grid(gridX,gridY))
    #get grid height
    gridWidth = length(gridX)
    #get grid width
    gridHeight = length(gridY)
  
    #plot the datapoints
    plot(
      data.frame(attribute1,attribute2), #create a data frame with the two selected columns
      col=ifelse(target==1,"blue","red"), #set the colors in wich to draw the points (based on target value)
      pch=20,cex=1.5,cex.lab=1.4 #visualization parameters
    )
    
    #get the predicted values for each element of the grid
    control = rpart.control(minucket = minucket)
    fit_minucket_2 <- rpart(trainDiabetes[,c(column1,column2)], data = trainDiabetes, method = 'class', control = control)
    pred = as.numeric(predict(fit_minucket_2, trainDiabetes, type = 'class'))
    #add the grid points to the plot
    points(grid,pch='.',col=ifelse(pred==1,"blue","red"))
    #add the decision boundary to the plot
    contour(gridX,gridY,matrix(pred,gridWidth,gridHeight),levels=1.5, add=TRUE, d=FALSE, lty=1)
  }
}
```
#depth
```R
sens_depth = vector(mode = 'list')
error_depth = data.frame(matrix(ncol = 3, nrow = 0))
```


```R
for (maxdepth in seq(2,6,1)) {
    control = rpart.control(maxdepth = maxdepth)
    fit_depth <- rpart(trainDiabetesY~., data = trainDiabetes, method = 'class', control = control)
    predict_unseen_depth <-predict(fit_depth, testDiabetes, type = 'class')
    table = table(testDiabetesY, predict_unseen_depth)
    metrics_tuned = misclassCounts(as.numeric(as.factor(predict_unseen_depth))-1,as.numeric(testDiabetesY)-1)
    sens_depth = append(sens_depth, metrics_tuned$metrics$Sens)
    trainDiabetesPred_depth = predict(fit_depth, trainDiabetes, type = 'class')
    testDiabetesPred_depth = predict(fit_depth, testDiabetes, type = 'class')
    trainError_depth = 1 - sum(trainDiabetesPred_depth==trainDiabetesY)/length(trainDiabetesY)
    testError_depth = 1 - sum(testDiabetesPred_depth==testDiabetesY)/length(testDiabetesY)
    error_depth = rbind(error_depth,data.frame(maxdepth,trainError_depth,testError_depth))
}
maxY_d = max(error_depth$trainError_depth, error_depth$testError_depth)
```


```R
values_depth = c(2:6)
```


```R
md <- do.call(rbind, Map(data.frame, maxdepth_values = values_depth, sensitivities = sens_depth))
```


```R
ggplot(md, aes(x = maxdepth_values, y = sensitivities)) + geom_line() + geom_point() + labs(x = 'Value of maxdepth', y = 'Sensitivity')
```
```R
#create the empty plot with some parameters
plot(error_depth[,1],main="Evolution of the train and test errors",xlab="Value of maxdepth",ylab="Error",ylim = c(0,maxY_d),type="n",xaxt="n")
#set the axis for the number of neighbours (k)
axis(1, 1:length(error_depth[,1]), error_depth[,1])
#plot the error for the training data
lines(error_depth[,2],type="b",col='red',pch=20)
#plot the error for the testing data
lines(error_depth[,3],type="b",col='blue',pch=20)
legend("bottomright",lty=1,col=c("red","blue"),legend = c("Train ", "Test "))
```

```R
ncols = length(trainDiabetes)
depth = 5

for(column1 in seq(1,ncols-1,1)){
  #select a diferent second column that has not been a first columns
  for(column2 in seq(column1+1,ncols,1)){
    #get the values from the first column
    attribute1 = trainDiabetes[column1]
    #get the values from the second column
    attribute2 = trainDiabetes[column2]
    #turn the target variable into a number (for visualization purposes)
    target = as.numeric(trainDiabetesY)
    
    #create points in the x axis
    gridX = seq(
      min(attribute1),max(attribute1),
      (max(attribute1)-min(attribute1))/100
    )
    #create points in the y axis
    gridY = seq(
      min(attribute2),max(attribute2),
      (max(attribute2)-min(attribute2))/100
    )
    #create a grid with the points
    grid = as.matrix(expand.grid(gridX,gridY))
    #get grid height
    gridWidth = length(gridX)
    #get grid width
    gridHeight = length(gridY)
  
    #plot the datapoints
    plot(
      data.frame(attribute1,attribute2), #create a data frame with the two selected columns
      col=ifelse(target==1,"blue","red"), #set the colors in wich to draw the points (based on target value)
      pch=20,cex=1.5,cex.lab=1.4 #visualization parameters
    )
    
    #get the predicted values for each element of the grid
    control = rpart.control(maxdepth = depth)
    fit_depth_2 <- rpart(trainDiabetes[,c(column1,column2)], data = trainDiabetes, method = 'class', control = control)
    pred = as.numeric(predict(fit_depth_2, trainDiabetes, type = 'class'))
    #add the grid points to the plot
    points(grid,pch='.',col=ifelse(pred==1,"blue","red"))
    #add the decision boundary to the plot
    contour(gridX,gridY,matrix(pred,gridWidth,gridHeight),levels=1.5, add=TRUE, d=FALSE, lty=1)
  }
}
```